---
title: 'Guided Prject: Predicting Car Prices'
author: "Michael Opiekun"
date: "5/28/2021"
output: html_document
---
# Predicting Car Prices using Machine Learning

Use k-nearest neighbors algorithm to predict car prices based on various characteristics including body style, engine type and horsepower..


```{r setup}
library(readr)
library(tibble)
library(dplyr)
library(tidyr)
library(caret)
```

## Data
Remove non-continuous data, make sure all columns are of numeric type, removes NAs from variable to predcit (price). 

1. symboling: -3, -2, -1, 0, 1, 2, 3. 
2. normalized-losses: continuous from 65 to 256. 
3. make: 
alfa-romero, audi, bmw, chevrolet, dodge, honda, 
isuzu, jaguar, mazda, mercedes-benz, mercury, 
mitsubishi, nissan, peugot, plymouth, porsche, 
renault, saab, subaru, toyota, volkswagen, volvo 

4. fuel-type: diesel, gas. 
5. aspiration: std, turbo. 
6. num-of-doors: four, two. 
7. body-style: hardtop, wagon, sedan, hatchback, convertible. 
8. drive-wheels: 4wd, fwd, rwd. 
9. engine-location: front, rear. 
10. wheel-base: continuous from 86.6 120.9. 
11. length: continuous from 141.1 to 208.1. 
12. width: continuous from 60.3 to 72.3. 
13. height: continuous from 47.8 to 59.8. 
14. curb-weight: continuous from 1488 to 4066. 
15. engine-type: dohc, dohcv, l, ohc, ohcf, ohcv, rotor. 
16. num-of-cylinders: eight, five, four, six, three, twelve, two. 
17. engine-size: continuous from 61 to 326. 
18. fuel-system: 1bbl, 2bbl, 4bbl, idi, mfi, mpfi, spdi, spfi. 
19. bore: continuous from 2.54 to 3.94. 
20. stroke: continuous from 2.07 to 4.17. 
21. compression-ratio: continuous from 7 to 23. 
22. horsepower: continuous from 48 to 288. 
23. peak-rpm: continuous from 4150 to 6600. 
24. city-mpg: continuous from 13 to 49. 
25. highway-mpg: continuous from 16 to 54. 
26. price: continuous from 5118 to 45400.

```{r Data}
cars <- read_csv('~/Documents/Programs_in_R/DataQuest/Data/imports-85.data', col_names = FALSE)
colnames(cars) <- c('symboling', 'normalized_losses', 'make', 'fuel_type', 'aspiration', 'num_of_doors', 'body_style', 'drive_wheels', 'engine_location', 'wheel_base', 'length', 'width', 'height', 'curb_weight', 'engine_type', 'num_cylinders','engine_size', 'fuel_system', 'bore', 'stroke', 'compression_ratio', 'horsepower', 'peak_rpm', 'city_mpg', 'highway_mpg', 'price')

# Remove all columns that are not usable in k-nearest neighbors (non numbers)
cars <- cars %>% select(-symboling,  -make, -fuel_type, -aspiration, -num_of_doors, -body_style, -drive_wheels, -engine_location, -engine_type, -num_cylinders, -fuel_system)
cars$normalized_losses <- as.numeric(as.character(cars$normalized_losses))
cars$bore <- as.numeric(as.character(cars$bore))
cars$stroke <- as.numeric(as.character(cars$stroke))
cars$horsepower <- as.numeric(as.character(cars$horsepower))
cars$peak_rpm <- as.numeric(as.character(cars$peak_rpm))
cars$price <- as.numeric(as.character(cars$price))

cars <- cars %>% filter(!is.na(price))
cars <- tibble(cars)
```

## Test relationships with price using featureplot()

```{r Featureplot}
featurePlot(x = cars[,1:15], y = cars$price) 

```
There seems to be a positive relationship between price and horsepower, curb-weight, engine_size, bore (weaker), wheel base, length, and width. There seems to be a negative relationship between price and city mpg and highway mpg.

## Create Data Partition of 80/20

```{r Partition}
training_indices <- createDataPartition(cars[['price']], p = .8, list =  FALSE)
training_cars <- cars[training_indices,]
test_cars <- cars[-training_indices,]

```


```{r traincontrol}
train_control <- trainControl(method = "cv", number = 4)
```

## Training the model
```{r Training}
knn_model <- train(price ~ highway_mpg + engine_size,
                    data = cars,
                    method = "knn",
                    trControl = train_control,
                    preProcess = c("center", "scale"))


```

## Predictions

```{r Predictions}
predictions <- predict(knn_model, newdata = test_cars)
postResample(pred = predictions, obs = test_cars$price)

```




