---
title: "Creating and Effective Data Analysis Workflow, part 2"
author: "Michael Opiekun"
date: "4/4/2021"
output: html_document
---
## Working to create an efficient data workflow by working through a Guided Data quest Project

# Observe Sales Data set
```{r Ingest Data and Library}
library(tidyverse)
library(tibble)
library(purrr)
library(lubridate)
library(stringr)

sales <- read.csv("~/Documents/Programs_in_R/DataQuest/Data/sales2019.csv")

dim(sales) #5000 by 5
colnames(sales) # "date" "user_submitted_review" "title" "total_purchased"                        "customer_type"      
#Types of columns
for(col in colnames(sales)) {
  paste0(col, " : ", typeof(sales[[col]])) %>% print
}

unique(sales) # there is missing data in total_purchased and user_submitted_review

# Is there missing data anywhere?
for (col in colnames(sales)) {
  paste0(col, 
         ", number of missing data rows: ", 
         is.na(sales[[col]]) %>% sum) %>% print
}

```

# Cleaning the Data

## Getting rid of missing values

```{r Getting Rid of Missing Values}

# Remove NA values from reviews 
complete_sales <-  sales %>% filter(
  !is.na(user_submitted_review)
)    #Down to 4544 rows


#Calculate mean for books purchased
mean_purchased <- complete_sales %>%
  filter(!is.na(total_purchased)) %>%
  pull(total_purchased) %>%
  mean

#Fill in missing values in total_purchased
complete_sales <- complete_sales %>%
  mutate(
    imputed_purchases = if_else(is.na(total_purchased), 
                                     mean_purchased,
                                as.numeric(total_purchased))
  )
```

# Determine if postive or negative review

## Classifying reviews

```{r Good Bad Review}
#Defining positive reviews
review_positive <- function(review) {case_when(
  str_detect(review,"a lot") ~ TRUE,
  str_detect(review,"Awesome") ~ TRUE,
  str_detect(review,"Good") ~ TRUE,
  str_detect(review,"Never") ~ TRUE,
  str_detect(review,"OK") ~ TRUE,
  TRUE~FALSE #none of the above words contained in the review
  )
}

complete_sales <- complete_sales %>% mutate(
  
  is_positive = unlist(map(user_submitted_review,review_positive))
)

```

# Was the new book program effective in increasing book sales?

## Compare books sold before and after program started

```{r Date Compared to before and after}
complete_sales <-  complete_sales %>% mutate(
  date_useable = if_else(mdy(date) < mdy("07/01/2019"), "Pre", "Post")
)

#Check if sales increased in period after
complete_sales %>% group_by(date_useable) %>% 
  summarize(sum(imputed_purchases))  
#Numbers decreased slightly after

#Check if individual books became more popular
complete_sales %>% group_by(date_useable, title) %>%
  summarize(sum(imputed_purchases)) %>% arrange(title, date_useable)
  

```

# Was customer type more likely to buy more books before or after program?

## Compare number of books per customer before and after program started

```{r Customer type before and after }
#Check if sales increased in period after based on customer type
complete_sales %>% group_by(date_useable, customer_type) %>% 
  summarize(sum(imputed_purchases)) %>% arrange(customer_type, date_useable)

#Business customers bought more but drop in individual sales


```

# Did review scores improve as a result of the program?

## Test review scores

```{r Review Scores}
complete_sales %>% group_by(date_useable, is_positive) %>% 
  summarize(sum(imputed_purchases)) %>% arrange(is_positive, date_useable)

# Positive reviews did increase after the start of the program but not by much.
```