---
title: 'Guided Project: Investigating Fandango Movie Ratings'
author: "Michael Opiekun"
date: "5/4/2021"
output: html_document
---

# Investigating Fandango Movie Ratings
Fandango movie ratings seem to be inlfated with rounding up of reviews rather than rounding with normal conventions. The goal is to see if any change in the rating system has been done since Walt Hickey found the discrepancy.

```{r setup Setup}
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
```

## Ingest data

```{r Data}
fandango_score_comparison <-  '~/Documents/Programs_in_R/DataQuest/Data/fandango_score_comparison.csv'
movie_ratings_16_17 <- '~/Documents/Programs_in_R/DataQuest/Data/movie_ratings_16_17.csv'

comparison <- read.csv(fandango_score_comparison)
movie_ratings <- read.csv(movie_ratings_16_17)

hickey <- comparison %>% select(FILM, Fandango_Stars, Fandango_Ratingvalue, Fandango_votes, Fandango_Difference)
ratings <- movie_ratings %>% select(movie, year, fandango)
```

Since ratings are not randomly selected, the goal is to see if ratings changed from 2016 to 2017. Check to see if both data sets include movies with more than 30 ratings (popular movies).
Fandango movies after Hickey ratings do not have number of fan votes.

Sample 10 movies and check reviews on Fandango site.

```{r Sample}
set.seed(1)
sample_n(ratings, size = 10)

```

Fandango now uses Rotten Tomato verified ratings since 2019

Reviews totals (review number not specified over 5000 reviews)
```{r Reviews}
set.seed(1)
sampled <- sample_n(ratings, size = 10)

reviews <- tibble(reviews = c(5001, 5001, 5001, 5001, 5001, 5001, 5001, 5001, 5001, 5001))
bind_cols(sampled, reviews)

```
 identifying popular movies from Hickey
```{r pop movies Hickey}
sum(hickey$Fandango_votes < 30) #(zero films with 30 or fewer reviews)

head(hickey$FILM, n = 10)

unique(ratings$year)

```
## Isolate movies releasd in 2015 and 2016 in seperate data sets

```{r 2015 2016}

hickey <- hickey %>% mutate(year = str_sub(FILM, -5, -2))

```
```{r}
table(hickey$year)
```
```{r 2015}
fandango_2015 <- hickey %>% 
  filter(year == 2015)
table(fandango_2015$year)

```
```{r 2016}
head(ratings)
table(ratings$year)
```

```{r}
fandango_2016 <- ratings %>% 
  filter(year == 2016)
table(fandango_2016$year)
```
## Compare Distributions 

First use Kernel Density Plots
```{r Kernel}
ggplot(data = fandango_2015, aes(x = Fandango_Stars)) +
  geom_density() +
  geom_density(data = fandango_2016, aes(x = fandango), color = "blue") + 
  labs(title = "Distributions for Fandango's ratings \n(2015 vs 2016)",
       x = "Stars",
       y= "Density") +
  scale_x_continuous(breaks = seq(0,5, by = .5), limits = c(0,5))



```

Left skewed for both plots but 2016 is skewed left of 2015 ratings. High general ratings given by reviewers.

Examine distributions and proportions

```{r Proportions}
fandango_2015 %>% 
  group_by(Fandango_Stars) %>% 
  summarize(Percentage = n() / nrow(fandango_2015) * 100)

fandango_2016 %>% 
  group_by(fandango) %>% 
  summarize(Percentage = n() / nrow(fandango_2016) * 100)

```
Average higher rating in 2015 than 2016

```{r mean and median}
mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
summary_2015 <- fandango_2015 %>% 
  summarize(year = "2015",
    mean = mean(Fandango_Stars),
    median = median(Fandango_Stars),
    mode = mode(Fandango_Stars))
summary_2016 <- fandango_2016 %>% 
  summarize(year = "2016",
            mean = mean(fandango),
            median = median(fandango),
            mode = mode(fandango))
# Combine 2015 & 2016 summary dataframes
summary_df <- bind_rows(summary_2015, summary_2016)
# Combine Dataframes
summary_df <- summary_df %>% 
  gather(key = "statistic", value = "value", - year)

summary_df
```
```{r plotting}
ggplot(data = summary_df, aes(x = statistic, y = value, fill = year)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparing Mean Median and Mode: 2015 vs 2016",
       x = "",
       y = "Stars")
```
Mean rating was lower in 2016 and fewer higher ratings.

```{r Mean Change}
means <- summary_df %>% 
  filter(statistic == "mean")
means %>% 
  summarize(change = (value[1] - value[2]) / value[1])
```
5% change in ratings. Therefore there was a change though slight after Hickey's review.