---
title: 'Guided Project: Analyzing Movie Ratings'
author: "Michael Opiekun"
date: "4/29/2021"
output: html_document
---

```{r setup}
library(dplyr)
library(rvest)
library(ggplot2)
```

This project is to engage in webscraping movie ratings to identify the top 30 movies between March and July 2020. Then check how they relate to usr correlations.

### Url
```{r Url}
url <- "http://dataquestio.github.io/web-scraping-pages/IMDb-DQgp.html"

wp_content <- read_html(url)
```

## AS numeric and Parse
```{r}
as.numeric("10.50")
as.numeric(c("14.59", "3.14", "55"))
readr::parse_number(c("14 min", "17,35", "(2012)", "1,2,3,4"))
stringr::str_trim(" Spaces before and after should disappear    ")

```

## Extract Movie Titles and Years
```{r Titles}
#Select Title
selector <- ".lister-item-header a"
# Extract title
titles <- wp_content %>%
  html_nodes(selector) %>%
  html_text()

titles
```
```{r Years}
#year selector
selector_2 <- ".lister-item-year"
# Extract Years
year <- wp_content %>%
  html_nodes(selector_2) %>%
  html_text()
# Parse Years from other data
years <- readr::parse_number(year)
years
```


## Extract Runtimes and Genres

```{r Runtimes and Genres}
#Select Runtime
selector_3 <- ".runtime"
#Extract Runtime
runtimes <- wp_content %>%
  html_nodes(selector_3) %>%
  html_text()
#Parse Runtime numbers
runtimes <-  readr::parse_number(runtimes)
runtimes

#Select Genres
selector_4 <- ".genre"
#Extract Genres
genre <- wp_content %>%
  html_nodes(selector_4) %>%
  html_text()
#Remove Whitespace from numbers
genre <-  stringr::str_trim(genre)
genre

```
## Selecting Attributes

```{r User Ratings and Metascores}
#Select Ratings
selector_5 <- ".ratings-imdb-rating"
attrtibute <- "data-value"
#Extract Ratings
user_ratings <- wp_content %>%
  html_nodes(selector_5) %>%
  html_attr(attrtibute)
#Parse Ratings numbers
user_ratings <-  readr::parse_number(user_ratings)
user_ratings

#Select Metascore
selector_6 <- ".ratings-metascore"
#Extract Metascore
metascore <- wp_content %>%
  html_nodes(selector_6) %>%
  html_text()
#Remove Whitespace from numbers
metascore <-  readr::parse_number(stringr::str_trim(metascore))
metascore

```
## Extract Votes

```{r Votes}
#Select Metascore
selector_7 <- ".sort-num_votes-visible :nth-child(2)"
#Extract Metascore
votes <- wp_content %>%
  html_nodes(selector_7) %>%
  html_text()
#Remove Whitespace from numbers
votes <-  readr::parse_number(stringr::str_trim(votes))
votes

```
## Deal with missing values
Add NA to missing values and remove 17 from vectors

```{r Missing Values}
append_vector <- function(vector, inserted_indices, values){

  ## Creating the current indices of the vector
  vector_current_indices <- 1:length(vector)

  ## Adding small amount of values (between 0 and 0.9) to the `inserted_indices`
  new_inserted_indices <- inserted_indices + seq(0, 0.9, length.out = length(inserted_indices))

  ## Appending the `new_inserted_indices` to the current vector indices
  indices <- c(vector_current_indices, new_inserted_indices)

  ## Ordering the indices
  ordered_indices <- order(indices)

  ## Appending the new value to the existing vector
  new_vector <- c(vector, values)

  ## Ordering the new vector wrt the ordered indices
  new_vector[ordered_indices]
}

#Use append_vector funtion to insert NA into metascores vector in psotions 1, 1, 1, 13, 24

metascore <- append_vector(metascore, c(1, 1, 1, 13, 24), NA)

#Remove 17th element from titles, years, runtimes, genres, and metascores
titles <- titles[-17]
years <- years[-17]
runtimes <- runtimes[-17]
genre <- genre[-17]
metascore <- metascore[-17]

```

## Create Dataframe with all values and plot

```{r dataframe and plot}
movies_df <- tibble( "titles" = titles, 
                     "years" = years, 
                     "runtimes" = runtimes, 
                     "genres" = genre, 
                     "user_ratings" = floor(user_ratings), 
                     "metascores" = metascore, 
                     "votes" = votes)

ggplot(data = movies_df, aes(x = user_ratings,
                             y = votes,
                             group = user_ratings)) +
  geom_boxplot()



```



















