---
title: 'Guided Project: Answering Business Questions using SQL'
author: "Michael Opiekun"
date: "4/17/2021"
output: html_document
---
USing SQL in R to answer business questions based on chinook database and data.
Looking at data for music sales.


```{r setup}
library(RSQLite)
library(DBI)
library(ggplot2)

```

Open ~/Documents/Programs_in_R/DataQuest/Data/chinook.db to view data

```{r Open Data}

db <- '~/Documents/Programs_in_R/DataQuest/Data/chinook.db'
run_query <- function(q) {
  conn  <- dbConnect(SQLite(), db)
  result <- dbGetQuery(conn, q)
  dbDisconnect(conn)
  return(result)
}

show_tables <-  function() {
  q = "SELECT name, type FROM sqlite_master WHERE type IN ('table','view');"
  return(run_query(q))
}

show_tables

```

# Recommend Signing Artist by Genre
Query to see number of tracks sold by genre in the US

```{r Tracks Sold Genre}
sold_genre_us = 'WITH usa_tracks_sold AS
                (
                SELECT il.* FROM invoice_line il
                INNER JOIN invoice inv ON il.invoice_id = inv.invoice_id
                INNER JOIN customer c ON inv.customer_id = c.customer_id
                WHERE c.country = "USA"
                )
     SELECT 
         g.name genre,
         COUNT(uts.invoice_line_id) total_sold,
         CAST(COUNT(uts.invoice_line_id) AS FLOAT)/(SELECT COUNT(*) FROM                          usa_tracks_sold) percent_sold
                FROM usa_tracks_sold uts
                INNER JOIN track t on t.track_id = uts.track_id
                INNER JOIN genre g on g.genre_id = t.genre_id
                GROUP BY 1
                Order By 2 DESC
                Limit 10;'
run_query(sold_genre_us)

# Plot using bar plot

genre_sales = run_query(sold_genre_us)
ggplot(data = genre_sales, aes(x = reorder(genre, -percent_sold),
                              y = percent_sold)) + geom_bar(stat = "identity")


```

Based on plot, Punk Pop and BLues are the top sold and therefore recommended are:
- Red Tone (Punk)
- Slim Jim Bites (Pop)
- Meteor and the Girls (Blues)


# Investigating Sales Associate Performance

```{r sales associate}
emp_sales_performance = 'WITH employee_rep_sales AS
              (
              SELECT c.customer_id,
              c.support_rep_id,
              SUM(i.total) total
              FROM customer c
              INNER JOIN invoice i on i.customer_id = c.customer_id
              GROUP BY 1, 2
              )
  SELECT e.first_name || " " || e.last_name employee_name,
         e.hire_date hire_date,
         SUM(emp.total) total_sales
         FROM employee_rep_sales emp
         INNER JOIN employee e ON e.employee_id = emp.support_rep_id
         GROUP BY 1;'
run_query(emp_sales_performance)

#Plotting results
employee_sales = run_query(emp_sales_performance)
ggplot(data = employee_sales, aes(x = reorder(employee_name, -total_sales),
                              y = total_sales)) + geom_bar(stat = "identity")
```
Jane Peacock sold most but also been working the longest. Standardization needed.


# Visualizing Sales By Country
Total number of customers
Total value of sales
Average value of sales per customer
Average order value

```{r Sales by Country}
sales_country = '
WITH country_other_sales AS
    (SELECT 
        CASE
          WHEN 
            (SELECT COUNT(*) FROM customer
            WHERE country = c.country
            ) = 1 THEN "Other"
            ELSE c.country
        END AS country,
        c.customer_id,
        il.*
    FROM invoice_line il
    INNER JOIN invoice i ON i.invoice_id = il.invoice_id
    INNER JOIN customer c ON c.customer_id = i.customer_id
    )
SELECT 
    country,
    customers,
    total_sales,
    avg_sales,
    avg_order
FROM
  (
    SELECT
      country,
      COUNT(distinct customer_id) customers,
      SUM(unit_price) total_sales,
      SUM(unit_price)/COUNT(distinct customer_id) avg_sales,
      SUM(unit_price)/COUNT(distinct invoice_id) avg_order,
      CASE
        WHEN country = "Other" THEN 1
        ELSE 0
      END AS sort
    FROM country_other_sales
    GROUP BY country
    ORDER BY sort, total_sales DESC
  );
'
run_query(sales_country)

# Visualizations for each country metric

country_sales = run_query(sales_country)

ggplot(data = country_sales, aes(x = reorder(country, -total_sales),
                                 y = total_sales,
                                 fill = country)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Total Sales by Country",
        aes(x = "Country",
            y = "Total Sales")

    ) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(data = country_sales, aes(x = reorder(country, -avg_sales),
                                 y = avg_sales,
                                 fill = country)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Sales by County",
        aes(x = "Country",
            y = "Average Sales")

    ) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(data = country_sales, aes(x = reorder(country, -customers),
                                 y = customers,
                                 fill = country)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Customers per Country",
        aes(x = "Country",
            y = "Customers")

    ) + theme(axis.text.x = element_text(angle = 45, hjust = 1))




```

US has highest sales and highest number of customers. Czech Republic has highest average sales.



# Albums vs Individual Tracks in Sales

```{r Albums vs Tracks}
albums_tracks = '
WITH invoice_first_song AS
    (
    SELECT 
      il.invoice_id invoice_id,
      MIN(il.track_id) first_track_id
    FROM invoice_line il
    GROUP BY 1
    )

SELECT 
  album_pur,
  COUNT(invoice_id) number_of_invoices,
  CAST(COUNT(invoice_id) AS FLOAT)/(
                                    SELECT COUNT(*) FROM invoice
                                  ) percent
FROM
  (
    SELECT
      ifs.*,
      CASE
        WHEN
          (
            SELECT t.track_id FROM track t
            WHERE t.album_id = (
                                SELECT t2.album_id FROM track t2
                                WHERE t2.track_id = ifs.first_track_id
                                )
            EXCEPT
            
            SELECT il2.track_id FROM invoice_line il2
            WHERE il2.invoice_id = ifs.invoice_id
          ) IS NULL
          
        AND
          (
            SELECT il2.track_id FROM invoice_line il2
            WHERE il2.invoice_id = ifs.invoice_id
            
            EXCEPT
            
            SELECT t.track_id FROM track t
            WHERE t.album_id = (
                                SELECT t2.album_id FROM track t2
                                WHERE t2.track_id = ifs.first_track_id
                                )
          ) IS NULL
          THEN "yes"
          ELSE "no"
      END AS "album_pur"
    FROM invoice_first_song ifs
  )
GROUP BY album_pur;
'

run_query(albums_tracks)




```

Albums account for almost 20% of sales. Continue to keep buying items.














