---
title: 'Guided Project: Investigating Covid-19 Virus Trends'
author: "Michael Opiekun"
date: "4/1/2021"
output: html_document
---
## Summary

This project is to deliniate trends in the Covid-19 Pandemic using real world data. The goal is to increase understanding of data analytics in R and oberserve trends among the World Covid-19 data.

# Organizatining the data

```{r opening and reading data}
# Open data from covid19.csv and observe structure and dim size
library(readr)
covid_df <- read_csv("~/Downloads/covid19.csv")
dim(covid_df)
vector_cols <- colnames(covid_df)
vector_cols
head(covid_df)

# Use glipse to identify data kinds
library(tibble)
glimpse(covid_df)

```

## Oragnizing Data

Organizing data to analyze country data

```{r cleaning and organizing data}
covid_df_all <- covid_df %>% filter(Province_State == "All States") # Filter for full country
covid_df_all_states <- covid_df_all[c(-5)] #Remove 'Province_State' column as the column is homogeneous
```

## Select Specific Data

```{r Data Selection}
covid_df_all_states_cumulative <- select(covid_df_all_states, Date, Continent_Name, Two_Letter_Country_Code, positive, hospitalized, recovered, death, total_tested)
covid_df_all_states_daily <- select(covid_df_all_states, Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)
```

# Identifying Highest Fatality Rates

## Summarizing Data Based on Continent_Name and Two_Letter_Country_Code

``` {r Summarizing Deaths}
library(dplyr)
covid_df_all_states_cumulative_max <- covid_df_all_states_cumulative %>%     group_by(Continent_Name, Two_Letter_Country_Code) %>% summarize(max = max(death) ) %>% filter(max>0)
covid_df_all_states_cumulative_max
```

## Displaying max number of deaths by country, colored by continent

```{r plotting deaths}
library(ggplot2)

qplot(x = Two_Letter_Country_Code,
      y = max,
      col = Continent_Name,
      data =  covid_df_all_states_cumulative_max
      )
```

##Conclusion: Answering question: Which countries have had the highest fatality rates?
```{r top 3 fatality}
death_top_3 <- c("US", "IT", "GB")
```

# Extracting the top ten cases countries data

## Summarizing data based on Country_Region column

```{r Summarizing for Top 10 Cases}
covid_df_all_states_daily_sum <-  covid_df_all_states_daily %>% group_by(Country_Region) %>% summarize(tested = sum(daily_tested),
                                       positive = sum(daily_positive),
                                       active = sum(active),
                                       hospitalized = sum(hospitalizedCurr)
                                       ) %>% arrange(-tested)
covid_df_all_states_daily_sum
```

## Extract top 10 results

```{r top 10}
covid_top_10 <-  head(covid_df_all_states_daily_sum,10)
```

# Identify the countires that have had the highest number of positive cases against the number of tests

## Create vectors from covid_top_10 dataframe

```{r top 10 vector creation}
countries <- covid_top_10$Country_Region
tested_cases <- covid_top_10$tested
positive_cases <- covid_top_10$positive
active_cases <- covid_top_10$active
hospitalized_cases <- covid_top_10$hospitalized
```

## Naming Vectors

```{r Naming top 10 vetors}
names(countries) <- countries
names(tested_cases) <- countries
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
```

## Identify top three positive against tested cases

```{r Top 10 positive v Tested}
positive_cases/tested_cases
```

## Conclusion

```{r top 3 positive v Tested}
positive_tested_top_3 <- head(positive_cases/tested_cases,3)
```

# Identifying countries affected by population

## Creating matrix

```{r Creating Matrix}
# Create matrix covid_mat
covid_mat <- cbind(tested_cases, positive_cases, active_cases, hospitalized_cases)

# Create population vector
population <- c(331002651, 145934462, 60461826, 1380004385, 84339067, 37742154, 67886011, 25499884, 32971854, 37846611)

# Dividing the matrix by the population vector
covid_mat <- covid_mat*100/population

covid_mat
```

## Ranking for every measurement in matrix covid_mat

```{r Ranking matrix}
tested_cases_rank <- rank(covid_mat[,"tested_cases"])
positive_cases_rank <- rank(covid_mat[,"positive_cases"])
active_cases_rank <- rank(covid_mat[,"active_cases"])
hospitalized_cases_rank <- rank(covid_mat[,"hospitalized_cases"])

# Combining rankings
covid_mat_rank <- rbind(tested_cases_rank, positive_cases_rank, active_cases_rank, hospitalized_cases_rank)

covid_mat_rank

covid_mat_rank[1,]

covid_mat_rank[-1,]

colSums(covid_mat_rank)
```

## Conclusion 
```{r Ranking Conclusion}
best_effor_tested_cased_top_3 <- c("India", "United Kingdom", "turkey")

most_affected_country <- "Italy"
least_affected_country <- "India"
```

# Culminating Everything

```{r Putting Everything Together}
question_list <- list(
  "Which countries have had the highest number of deaths due to COVID-19?",
  "Which countries have had the highest number of positive cases against the number of tests?",
  "Which countries have made the best effort in terms of the number of COVID-19 tests conducted related to their population?",
  "Which countries were ultimately the most and least affected related to their population?"
  )

answers_list <- list(
  "Death" = death_top_3,
  "Positive tested cases" = positive_tested_top_3,
  "The best effort in test related to the population" = best_effor_tested_cased_top_3,
  "The most affected country related to its population" = most_affected_country,
  "The least affected country related to its population" = least_affected_country
)

datasets = list(original = covid_df, all_states = covid_df_all_states, cumulative = covid_df_all_states_cumulative, daily = covid_df_all_states_daily
)

matricies <- list(covid_mat, covid_mat_rank)

vectors <- list(vector_cols, population, countries)

data_structure_list <- list("data_frame" = datasets, "matrix"= matricies, "vector" = vectors)

covid_analysis_list <- c(question_list, answers_list, data_structure_list)
```