---
title: "Guided Project - Predicting Condominium Sale Prices"
author: "Michael Opiekun"
date: "5/24/2021"
output: html_document
---
## Compare Condominiums in New York City

Identify how well the size of a condominium explains or predicts the sale price across NYC as a whole.

How well does the size of a condo explain the sale price in each individual borough.

```{r setup}
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(broom)
library(stringr)
library(magrittr)
library(purrr)
```

## Data
```{r Data}
#Read in data
NYC_property_sales <- read_csv('~/Documents/Programs_in_R/DataQuest/Data/NYC_property_sales.csv')
sort(table(NYC_property_sales$building_class_at_present))

# convert colnames to lower case
colnames(NYC_property_sales) %>%str_replace_all("\\s", "_") %>% tolower()

# Convert to title case
NYC_property_sales <- NYC_property_sales %>% mutate(neighborhood = str_to_title(neighborhood)) %>% 
  mutate(building_class_category = str_to_title(building_class_category)) %>% 
  mutate(address = str_to_title(address))

# Remove unnecessary columns and only take distinct rows
NYC_property_sales <- NYC_property_sales %>% select(-apartment_number) %>% distinct()

# Filter data for properties above $10000, zero gross square footage, and drop NAs
NYC_property_sales <- NYC_property_sales %>% filter(sale_price > 10000) %>%
  filter(gross_square_feet > 0) %>%
  drop_na(c(gross_square_feet, sale_price))

# Arrange alphabetically by borough and neighborhood
NYC_property_sales <- NYC_property_sales %>% arrange(borough, neighborhood)

# Save results to csv file for future use
# write_csv(NYC_property_sales, "NYC_property_sales.csv")
```
## Generate scatter plots to visualize sale_price versus gross_square_feet to explore bivaritate relationships to examine direction, linearity, and strength.
```{r Looking for Relationships}
NYC_condos <- NYC_property_sales %>% filter(building_class_at_time_of_sale == 'R4')
```
```{r Plot}

ggplot(data = NYC_condos, aes(x = gross_square_feet, y = sale_price, color = borough)) + geom_point(alpha = .3) + 
  scale_y_continuous(labels = scales::comma, limits = c(0, 100000000)) + xlim(c(0, 12000)) + geom_smooth(method = 'lm', se = FALSE) + theme_gray() +
  labs(title = "Sale Price Compared to Square Footage of Condos",
       x = "Size (Gross Square Feet)",
       y = "Sale Price")


# Zoom in
ggplot(data = NYC_condos, aes(x = gross_square_feet, y = sale_price, color = borough)) + geom_point(alpha = .3) + 
  scale_y_continuous(labels = scales::comma, limits = c(0, 20000000)) + xlim(c(0, 6000)) + geom_smooth(method = 'lm', se = FALSE) + theme_gray() +
  labs(title = "Sale Price Compared to Square Footage of Condos",
       x = "Size (Gross Square Feet)",
       y = "Sale Price")

```
```{r Facet Wrap}
ggplot(data = NYC_condos, aes(x = gross_square_feet, y = sale_price)) + geom_point(alpha = .3) + 
  facet_wrap(~ borough, scales = "free", ncol = 2) +
  scale_y_continuous(labels = scales::comma) +
  geom_smooth(method = 'lm', se = FALSE) + theme_gray() +
  labs(title = "Sale Price Compared to Square Footage of Condos",
       x = "Size (Gross Square Feet)",
       y = "Sale Price")
```

## Identify Outliers

```{r Outliers}
NYC_condos %>% arrange(desc(sale_price)) %>% head
```


Remove multi-unit sales including 165 East 66th St, Resi (full block sale)

```{r}
# Copy the dataframe
NYC_condos_original <- NYC_condos

# Remove 165 East 66th St, Resi
NYC_condos <- NYC_condos %>% filter(address != "165 East 66th St, Resi")

```

Will leave the most expensive sale since it is legitimate

## More than one sale
```{r}
NYC_condos %>% 
  filter(borough == "Brooklyn") %>% 
  arrange(desc(sale_price))
```

Large number of units sold at same price on the same day, will remove since units were all sold for lump sum and not individual prices

```{r Multi-unit}
multi_unit_sales <- NYC_condos %>% 
  group_by(sale_price, sale_date) %>% 
  filter(n() >= 3) %>% 
  arrange(desc(sale_price))
```
Includes mostly multi-unit sales and can be removed with minor consequences.

```{r Remove multi-unit purchases}

NYC_condos <- NYC_condos %>%
  group_by(sale_price, sale_date) %>%
  filter(n() <= 2) %>%
  ungroup()

```

## Modeling Relationships

```{r}
# sale_price explained by gross_square_feet for the NYC_condos
NYC_condos_lm <- lm(sale_price~gross_square_feet, data = NYC_condos)
summary(NYC_condos_lm)

# sale_price explained by gross_square_feet for the NYC_condos_original
NYC_condos_original_lm <- lm(sale_price~gross_square_feet, data = NYC_condos_original)
summary(NYC_condos_original_lm)
```

```{r}
confint(NYC_condos_lm)
confint(NYC_condos_original_lm)
sigma(NYC_condos_lm)
sigma(NYC_condos_original_lm)
```

Two linear models were tested to see if there was a relationship between the sale price and the gross square footage of condos in NYC. First the data were tested without multi-unit sales and then with multi-unit sales. Both models showed a strong, positive relationship described by a high t-statistic and low p-value, between sale price and square footage but the t-statistic was higher in the data with multi-unit sales removed.
There is a lower confidence interval for the original data is lower than the and most likely attributed to removal of high dollar multi-unit sales. The RSE (residual standard error) is lower after removing the multi-unit sales. R^2 for modified data is also double that of the original data meaning the sales are better explained by square footage.

```{r}
ggplot(data = NYC_condos, aes(x = gross_square_feet, y = sale_price)) + geom_point(alpha = .3) + 
  facet_wrap(~ borough, scales = "free", ncol = 2) +
  scale_y_continuous(labels = scales::comma) +
  geom_smooth(method = 'lm', se = FALSE) + theme_gray() +
  labs(title = "Sale Price Compared to Square Footage of Condos",
       x = "Size (Gross Square Feet)",
       y = "Sale Price")
```

Very few outliers remain for each borough that changes the graphs disproportionately.The sale price seems to be very well related to the square footage in each borough after removing multi-unit sales.

## Nest Data and follow Broom workflow

1. Nest a dataframe by a categorical variable with the nest() function from the tidyr package - we will nest by borough.
2. Fit models to nested dataframes with the map() function from the purrr package.
3. Apply the broom functions tidy(), augment(), and/or glance() using each nested model - we'll work with tidy() first.
4. Return a tidy dataframe with the unnest() function - this allows us to see the results.

```{r Nesting}
NYC_nested <- NYC_condos %>% group_by(borough) %>% nest()
```
```{r}
print(NYC_nested)
```

Apply lm function to each borough
```{r Map lm}
NYC_coefficients <- NYC_nested %>% mutate(lm =  map(.x = data, 
                                                    .f = ~lm(sale_price ~ gross_square_feet, 
                                                             data = .)))
```
```{r}
#Inspect data structure
print(NYC_coefficients)

# View results for Bronx
summary(NYC_coefficients$lm[[1]])
```
## Generate tidy dataframe
```{r Tidy}
NYC_coefficients <- NYC_coefficients %>%
  mutate(tidy_coefficients = map(.x = lm, 
                                 .f = tidy, 
                                 conf.int = TRUE))
NYC_coefficients
print(NYC_coefficients$tidy_coefficients[[2]])
```

## Unnest Tidy data frame 

```{r Unnest}
NYC_coefficients_tidy <- NYC_coefficients %>%
  select(borough, tidy_coefficients) %>%
  unnest(cols = tidy_coefficients)
print(NYC_coefficients_tidy)
```

## Filter slope
```{r Slope}
NYC_slope <- NYC_coefficients_tidy %>%
  filter(term == "gross_square_feet") %>%
  arrange(estimate)
NYC_slope
```
Low p value and high t-statistic expresses a relationship with between sale price and gross square footage. Manhattan has the highest price per foot and Staten Island has the lowest.

## Broom workflow on NYC_condos

```{r Broom NYC Condos}
NYC_summary <- NYC_condos %>% 
  group_by(borough) %>% 
  nest() %>% 
  mutate(lm =  map(.x = data, 
                   .f = ~lm(sale_price ~ gross_square_feet,
                            data = .))) %>%
  mutate(tidy_summary = map(.x = lm, 
                            .f = glance))
print(NYC_summary$tidy_summary[[2]])
```

## Unnest Single tidy dataframe includes regression summary statistics for each borough

```{r Unnest Glance}
NYC_summary_tidy <- NYC_summary %>%
  select(borough, tidy_summary) %>%
  unnest(cols = tidy_summary) %>%
  arrange(r.squared)
print(NYC_summary_tidy)
```

## Conclusions

When multi-unit sales were removed from the analysis, gross-square feet was found to be a good predictor of sale price in each borough. Manhattan was the best predicted (r^2 .63) and Queens was the worst predicted (r^2 .34). Certain boroughs had much higher price per gross footage compared to other boroughs.








