---
title: 'Guided Project: Analyzing Forest Fire Data'
author: "Michael Opiekun"
date: "4/7/2021"
output: html_document
---
## Observing forest fire data from Europe to better understand plotting in R

```{r Ingest Data}
library(tidyverse)

forest_fires <- read.csv("~/Documents/Programs_in_R/DataQuest/Data/forestfires.csv")

# each row represents a fire that occurred
forest_fires
```
## The indivdual column descriptions:
X: X-axis spatial coordinate within the Montesinho park map: 1 to 9
Y: Y-axis spatial coordinate within the Montesinho park map: 2 to 9
month: Month of the year: 'jan' to 'dec'
day: Day of the week: 'mon' to 'sun'
FFMC: Fine Fuel Moisture Code index from the FWI system: 18.7 to 96.20
DMC: Duff Moisture Code index from the FWI system: 1.1 to 291.3
DC: Drought Code index from the FWI system: 7.9 to 860.6
ISI: Initial Spread Index from the FWI system: 0.0 to 56.10
temp: Temperature in Celsius degrees: 2.2 to 33.30
RH: Relative humidity in percentage: 15.0 to 100
wind: Wind speed in km/h: 0.40 to 9.40
rain: Outside rain in mm/m2 : 0.0 to 6.4
area: The burned area of the forest (in ha): 0.00 to 1090.84

## Convert the month and day variables into a categorical variables

```{r convert months and days to categorical variables}
forest_fires %>% pull(month) %>% unique

forest_fires %>% pull(day) %>% unique

# define order of months and days
month_order <- c("jan","feb","mar","apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

#Assume Sunday is first day of the week
day_order <- c("sun","mon","tue","wed","thu","fri","sat")

forest_fires <-  forest_fires %>% mutate(month = factor(month, levels = month_order),
                                        day = factor(day, levels = day_order)
)

```


## Which months do forest fires happen the most? Which days of the week do forest fires happen the most?

```{r Most Fires by month and Day}
#Tibble: Fires by Month
forest_fires_month <- forest_fires %>% group_by(month) %>% 
                      summarize(fires_month = n())
#Tibble: Fires by Day
forest_fires_day <- forest_fires %>% group_by(day) %>% 
                      summarize(fires_day = n())
library(ggplot2)
#Visualize Fires by month
forest_fires_month %>% ggplot(aes(x = month, y = fires_month)) + geom_col() +
  labs(title = "Fire Frequency by Month",
       x = "Month",
       y = "Fires")

#Visualize Fires by month
forest_fires_day %>% ggplot(aes(x = day, y = fires_day)) + geom_col() +
  labs(title = "Fire Frequency by Day",
       x = "Month",
       y = "Fires")

```

Fires seem to occur most frequently in August and September and most on Friday, Saturday, and Sunday.

## Look at the relationships between other variables and frequencies of fores in August and September

```{r Other Factors relationship with Fires}

# Make tibble long
forest_fires_long <- forest_fires %>% pivot_longer(
  cols = c("FFMC","DMC","DC","ISI","temp","RH","wind","rain"
  ),
  names_to = "data_col",
  values_to = "values"
  
)

forest_fires_long %>% ggplot(aes(x = month, y = values)) + geom_boxplot() +
  facet_wrap(facets = vars(data_col),
             scales = "free_y") +
  labs(
    title = "Variable changes per month",
    x = "month",
    y = "Variable Value"
  )

```

## Look at Area of fire burned vs other variables

```{r Area vs Variables}

#Plot scatter plot for each variable
forest_fires_long %>% ggplot(aes(x = values, y = area)) + geom_point() +
  facet_wrap(facets = vars(data_col),
            scales = "free_x") +
  labs(
    title = "Variables Effect of Fire Area",
       x = "Variable Value",
       y = "Area (hectare)")

#some extreme outliers in each plot crushes the visualization

```

## Outlier Issues

Two points are extremely large burn areas that kill the plots. Removing them will help visualize better

```{r Outlier}

#Plot scatter plot for each variable without outliers
forest_fires_long %>% filter(area < 600) %>% ggplot(aes(x = values, y = area)) + geom_point() +
  facet_wrap(facets = vars(data_col),
            scales = "free_x") +
  labs(
    title = "Variables Effect of Fire Area",
       x = "Variable Value",
       y = "Area (hectare)")

```
Forest fires seem to be negatively correlated with ISI and rain and postively correlated with FFMC and DC.