---
title: 'Guided Project: Mobile App for Lottery Addiction'
author: "Michael Opiekun"
date: "5/10/2021"
output: html_document
---

```{r setup}
library(dplyr)
library(ggplot2)
library(tidyverse)
```

## Help build app that helps gambling addicts understand and estimate their odds in winning to help them realize that increasing odds of winning is difficult.

Questions:
What is the probability of winning the big prize with a single ticket?
What is the probability of winning the big prize if we play 40 different tickets (or any other number)?
What is the probability of having at least five (or four, or three, or two) winning numbers on a single ticket?

Create a function for factorial and combinations
```{r Fucntions}
factorial_1 <- function(n) {
  product <- 1
  for (i in 1:n) {
    product <- product * i
  }
  return(product)
}

combinations <- function(n,k) {
  combo <-factorial_1(n)/(factorial_1(k)*factorial_1(n-k))
  return(combo)
}

```

Write a function that takes in 6 values from 1-49 as a vector and gives the probability of winning.

```{r Probability of a ticket}
one_ticket_probability <- function(ticket) {
  total_combonations <- combinations(49,6)
  probability <- 1/total_combonations
  print_prob <- sprintf("%1.9f", probability)
  to_print <- paste("You have a ", print_prob,"% chance of winning the large prize.", sep = "")
return(to_print)
}

one_ticket_probability(c(4,5,6,7,8,9))

```

Use the fact that no matter what numbers are chosen, the probability of one ticket or combination of numbers does not change.

```{r Data}
lottery <- read.csv('~/Documents/Programs_in_R/DataQuest/Data/649.csv')

dim(lottery)
head(lottery)
```


```{r Practice for pmap}
data1 <- c(1, 3, 5)
data2 <- c(2, 4, 6)
data3 <- c(8, 9, 7)
unnamed_list <- list(data1, data2, data3)
first_vector <- unnamed_list[[1]]
named_list <- list('first' = data1, 'second' = data2, 'third' = data3)
first_item_sum <- named_list$first[1] + named_list$second[1] + named_list$third[1]

```

```{r pmap}
data <- list(data1, data2, data3)
averages <- pmap(data, function(x,y,z) {(x + y + z)/3})
first_average <- unlist(averages[1])


```

Applying function to vector from multiple rows
```{r Extract from Rows}
historical_nums <- pmap(list(a <- lottery$`NUMBER.DRAWN.1`,
                        b <- lottery$`NUMBER.DRAWN.2`,
                        c <- lottery$`NUMBER.DRAWN.3`,
                        d <- lottery$`NUMBER.DRAWN.4`,
                        e <- lottery$`NUMBER.DRAWN.5`,
                        f <- lottery$`NUMBER.DRAWN.6`),
                        function(a,b,c,d,e,f) {c(a,b,c,d,e,f)})

```

```{r Function}
check_historical_occurrences <- function(numbers, historic_numbers = historical_nums) {
  historical_win <- map(historic_numbers, function(x) {setequal(x,numbers)})
  past_matches <- sum(unlist(historical_win))
  statement <- paste("The numbers you entered have won ",
                     past_matches,
                     " times in the past. ",
                     "Your chance of winning the large prize in the next drawing is 0.0000072%.", sep = "")
  return(statement)
}


check_historical_occurrences(c(3,12,11,14,41,43))
check_historical_occurrences(c(8,9,10,11,12,13))

```
Count the times TRUE appears in matches to identify the number a set of number has won in the past.

Write function for the probability of winning with multiple tickets

```{r Multiple Tickets}
multi_ticket_probability <- function(number_of_tickets) {
  combinations_total <- combinations(49,6)
  win_probability <- number_of_tickets/combinations_total*100
  print_prob <- sprintf("%1.9f", win_probability)
  to_print <- paste("You have a ", print_prob, "% chance of winning the highest prize with ", number_of_tickets, " tickets purchased.", sep = "")
  return(to_print)
}

test_cases <- c(1, 10, 100, 10000, 1000000, 6991908, 13983816)

for (i in test_cases) {
  print(paste("For ", i, " tickets, ", multi_ticket_probability(i), sep = ""))
}


```

Including combination and then printing based on number of tickets purchased.


## Function for numbers to demonstrate probability of winning depending on the value of the integer in range from 3-5 for other winning options than the full six number option.

Creating a combination of a combination.

```{r Winning with fewer than 6 numbers}
probability_less_6 <- function(number) {
  ticket_combinations <- combinations(6,number)
  combinations_remaining <- combinations(49-6,6-number)
  outcomes <- ticket_combinations * combinations_remaining
  total_combinations <- combinations(49,6)
  
  probability <- (outcomes/total_combinations) *100
  
  for_print <- sprintf("%1.9f", probability)
  
  to_print <- paste("There is a ", for_print, "% chance of winning the prize for the ", number, " prize.", sep = "")
  return(to_print)
}

```

```{r test}
nums <- c(3,4,5)

for (i in nums) {
print(paste( "For ", i, " tickets, the proability of winning is ", probability_less_6(i), sep = ""))
}


```

