---
title: 'Guided Project: Creating An Efficient Data Analysis Workflow'
author: "Michael Opiekun"
date: "4/1/2021"
output: html_document
---

## Working through Creating an Efficient Data Analysis Workflow

# Observing Data

```{r Library}
library(tidyverse)
library(dplyr)
reviews <- read_csv("~/Documents/Programs_in_R/DataQuest/Data/book_reviews.csv")
dim(reviews)
col_names <- for (i in colnames(reviews)) {
  typeof(reviews[[i]])
}

#Unique Reviews
for (i in colnames(reviews)) {
  print("Unique values in column:")
  print(i)
  print(unique(reviews[[i]]))
  print("")
}
```

## Cleaning Data

```{r Cleaning}

reviews_cleaned <- reviews %>% filter(!is.na(review))#Keep only full reviews
dim(reviews_cleaned)
#206 reviews were removed for incompleteness
```

## Changing states to a consistent name

```{r State names}
reviews_cleaned <- reviews_cleaned %>% mutate(
  state = case_when(
    state == "California" ~ "CA",
    state == "New York" ~ "NY",
    state == "Florida" ~ "FL",
    state == "Texas" ~ "TX",
    TRUE ~ state # Stays the same if just postal code
    )
)
```

## Changing Review to numeral form

```{r Review to Review_num}
review_num <- reviews_cleaned %>% mutate(
  review_num = case_when(
    review == "Poor" ~ 1,
    review == "Fair" ~ 2,
    review == "Good" ~ 3,
    review == "Great" ~ 4,
    review == "Excellent" ~ 5
    
  ),
  is_high_review = if_else(review_num >= 4, TRUE, FALSE)
)

```

# Looking for most Profitable books

##Choosing Most Profitable based on books sold * price

```{r Most Profitable}
for_profitable <- reviews_cleaned %>% 
  group_by(book) %>% 
  summarize(
    purchased = n(),
    price = mean(price)
    ) %>%  arrange(-purchased)
most_profitable <- for_profitable %>% group_by(book) %>%
mutate(
  revenue = purchased*price 
) %>% arrange(-revenue)


```


Secrets of R for Advanced Students created the most revenue while Fundamentals of R for Beginners sold most copies.


The point of this project was to create a smooth workflow and determine most profitable book from R class books.







